---
title: "FItBit-Analysis"
author: "Satyendra Singh"
date: '2022-05-18'
output: html_document
---
##ANALYISIS ON FITBIT FITNESS TRACKER DATA

#OBJECTIVE-
HOW CAN A WELLNESS TECHNOLOGY COMPANY(Bellabeat) PLAY IT SMART ?

##IMPORTING LIBRARIES
```{r}
library(ggplot2)
```


```{r}
library(dplyr)
```


```{r}
#install.packages("skimr") for generating summary of the dataset 
library(skimr)
```


```{r}
#install.packages("janitor") for cleaning the dirty dataset 
library(janitor)
```


```{r}
library(tidyverse)
```
```{r}
library(lubridate)
```


```{r}
library(stringr)
```
#Importing the required dataset 
```{r}
daily_activity<-read_csv("D:/google cert/case_2/dailyActivity_merged.csv")
str(daily_activity)
```

```{r}
daily_sleep<-read_csv("D:/google cert/case_2/sleepDay_merged.csv")
str(daily_sleep)
```
```{r}
hourlyintensities<-read_csv("D:/google cert/case_2/hourlyIntensities_merged.csv")
str(hourlyintensities)
```
```{r}
Metsperminute<-read_csv("D:/google cert/case_2/minuteMETsNarrow_merged.csv")
str(Metsperminute)
```
```{r}
weightloginfo<-read_csv("D:/google cert/case_2/weightLogInfo_merged.csv")
str(weightloginfo)
```
##Removing Duplicates 
```{r}
daily_activity %>% distinct(Id)
daily_sleep %>% distinct(Id)
hourlyintensities %>% distinct(Id)
Metsperminute %>% distinct(Id)
weightloginfo %>% distinct(Id)
```
#Making new dataFrame out of the given dataset

##Sorting And FIltering
```{r}
Daily_sleep<-daily_sleep %>% distinct(Id,SleepDay,.keep_all = TRUE)
str(Daily_sleep)
```
```{r}
Weightloginfo<-weightloginfo %>% distinct(Id,.keep_all = TRUE)
str(Weightloginfo)
```
#copying data set
```{r}
Daily_activity<-daily_activity
```

#changing format of data using lubridate
```{r}
Daily_activity$ActivityDate1<-as.Date(Daily_activity$ActivityDate,format="%m/%d/%Y")
str(Daily_activity)
```
#fetching weekdays 
```{r}
Daily_activity$DayName<-weekdays(Daily_activity$ActivityDate1)
str(Daily_activity)
```
#Taking only columns that are required
```{r}
Activitylog<-Daily_activity %>%  select(Id,ActivityDate,DayName,TotalSteps,Calories,SedentaryMinutes)
str(Activitylog)
```
#sleep log
```{r}
sleeplog<-Daily_sleep %>% select(Id,SleepDay,TotalTimeInBed,TotalMinutesAsleep)
str(sleeplog)
```
```{r}
weightlog<-Weightloginfo %>% select(Id,BMI)
str(weightlog)
```
#Merging data
```{r}
MetPerMinutes1<-separate(Metsperminute,ActivityMinute,into=c("Date","Time"),sep=" ")
str(MetPerMinutes1)
```
#
```{r}
MetLogs<-MetPerMinutes1 %>%
  group_by(Id,Date) %>% 
  drop_na() %>% 
  summarise(AvgMETsparday=mean(METs))
str(MetLogs)
```
##Renaming the required columns
```{r}
colnames(Activitylog)[colnames(Activitylog)=="ActivityDate"]<-"Date"
colnames(sleeplog)[colnames(sleeplog)=="SleepDay"]<-"Date1"
sleeplog$Date<-strtrim(sleeplog$Date1,9)
sleeplog$Date1<-NULL
str(sleeplog)
```
```{r}
str(Activitylog)
```
#Merging the dataset and required columns
```{r}
mergedata1<-merge(Activitylog,sleeplog,by=c('Id',"Date"))
str(mergedata1)
```
```{r}
mergedata2<-merge(mergedata1,MetLogs,by=c("Id","Date"))
str(mergedata2)
```
##changing units and values in variables
#changing minutes to hour
```{r}
mergedata2$Sedentaryhrs<-mergedata2$SedentaryMinutes/60
mergedata2$SedentaryMinutes<-NULL #removing column
```
```{r}
mergedata2$TotalsleepHrs<-mergedata2$TotalMinutesAsleep/60
mergedata2$TotalMinutesAsleep<-NULL
```
```{r}
mergedata2$TotalTimeInBedHrs<-mergedata2$TotalTimeInBed/60
mergedata2$TotalTimeInBed<-NULL
```
```{r}
str(mergedata2)
```
#Merge for weight column
```{r}
#merge weights
mergeWt<-merge(mergedata2,weightlog,by="Id")
str(mergeWt)
```
#Extracting the required info from string
```{r}
#dataset
hourlyintensities
#active hour contains string 
```
```{r}
#Extracting time 
hourlyintensities$Time<-substr(hourlyintensities$ActivityHour,start = 11,stop = 21)
```
```{r}
#Extracting Date
hourlyintensities$Date<-substr(hourlyintensities$ActivityHour,1,9)
#changing format of date
hourlyintensities$Date1<-as.Date(hourlyintensities$Date,format = ("%m/%d/%Y"))
```
```{r}
#Extracting Day Names
hourlyintensities$DayName<-weekdays(hourlyintensities$Date1)
```
```{r}
str(hourlyintensities)
```
#VIZUALISATION of dataset
##VIZUALIZATION-1
```{r}
#VIZ.1:Total Steps on each day of the week
ggplot(mergedata2)+geom_point(mapping=aes(x=DayName,y=TotalSteps),color="Purple")+
  scale_x_discrete(limits=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))+
  labs(title = "Total Steps On Each Day of the Week",
       subtitle = "From Data of 24 Fitbit users")

```
##VIZ.2: Weekly steps taking pattern in a group of people who takes more than 8000 daily steps.
```{r}
mergedata2 %>% 
  filter(TotalSteps>8000) %>% 
  ggplot(mapping=aes(x=DayName,y=TotalSteps,fill=DayName))+
  geom_bar(stat="summary",fun="mean")+
  scale_x_discrete(limits=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))+
  labs(title="Avg Total Steps on each day of the week",
       subtitle = "among the people takes more then 8000 daily steps",
       y="Avg Total Steps")+
  theme(legend.position="none")
```
##VIZ.3: Steps pattern among those who takes less than 8000 daily steps.
```{r}
theme_set(theme_bw())
mergedata2%>% 
  filter(TotalSteps<8000) %>% 
  ggplot(mapping=aes(x=DayName,y=TotalSteps,fill=DayName))+
  geom_bar(stat="summary",fun="mean")+
  scale_x_discrete(limits=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))+
  labs(title="Avg Total Steps on each day of the week",
       subtitle = "among the people takes less then 8000 daily steps",
       y="Avg Total Steps")+
  theme(legend.position="none")
```
##VIZ.4: Steps pattern among those who takes less than 5000 daily steps.
```{r}
mergedata2%>% 
  filter(TotalSteps<5000) %>% 
  ggplot(mapping=aes(x=DayName,y=TotalSteps,fill=DayName))+
  geom_bar(stat="summary",fun="mean")+
  scale_x_discrete(limits=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))+
  labs(title="Avg Total Steps on each day of the week",
       subtitle = "among the people takes less then 5000 daily steps",
       y="Avg Total Steps")+
  theme(legend.position="none")
```

##VIZ.5: Users consistency in steps and their BMI with 22 days Data of 6 Fitbit Users.
```{r}
mergeWt %>% 
  ggplot(aes(DayName,TotalSteps,color=BMI,size=Date))+
  geom_point(alpha=0.5)+
  scale_x_discrete(limits = c("Monday" , "Tuesday", "Wednesday",
                              "Thursday" , "Friday" , "Saturday" , "Sunday"))+
  facet_wrap(~Id)+
  scale_color_viridis_b(option = "C")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Users consistancy in steps and their BMI",
       subtitle = " 22 days Data of 6 Fitbit Users")
```

##VIZ.6: Users consistency in steps and their Sleep Pattern.
```{r}
ggplot(data=mergedata2) +
  geom_point(mapping = aes(x=TotalsleepHrs,y=TotalSteps,
                           color=DayName),size=3) +
  labs(title = "Users consistancy in steps and their Sleep Pattern",
       subtitle = " 22 days Data of 24 Fitbit Users")
```
##VIZ.7: 6 Users consistency in steps and their Sleep Pattern in 22 days time oeriod.
```{r}
mergeWt %>% 
  ggplot(aes(DayName,TotalSteps,color=TotalsleepHrs))+
  geom_point(size=3)+
  scale_x_discrete(limits = c("Monday" , "Tuesday", "Wednesday",
                              "Thursday" , "Friday" , "Saturday" , "Sunday"))+
  facet_wrap(~Id)+
  scale_color_viridis_b(option = "C")+
  theme(axis.text.x = element_text(angle = 25))+
  labs(title = "Users consistancy in steps and their Sleep Pattern",
       subtitle = " 22 days Data of 6 Fitbit Users")

```
##VIZ.8: Total bed time and total sleep time.
```{r}
ggplot(data=mergedata2) +
  geom_jitter(mapping = aes(x=TotalTimeInBedHrs,y=TotalsleepHrs,color=DayName)) +
  labs(title = "Total bed time and total sleep time On Each Day of the Week", 
       subtitle = "From Data of 24 Fitbit users of 22 Days")+
  annotate("text",x=9,y=4,label="Less Sleeping Set",color="purple",
           fontface="bold",angle=30,size=4)
```
##VIZ.9: Relation between users daily calories burned and avg METs.
```{r}
ggplot(data=mergedata2) +
  geom_point(mapping = aes(x=Calories,y=AvgMETsparday,color=DayName)) +
  theme(legend.position = "none") +
  labs(title = "Relation between users daily calories burned and avg METs",
       subtitle = " 22 Days Data of 24 Fitbit users ",x="Daily Calories Burned")
```
##VIZ.10: Relation between users daily calories burned and Daily Sedentary Hrs.
```{r}
ggplot(data=mergedata2) +
  geom_smooth(mapping = aes(x=Sedentaryhrs ,y=Calories)) +
  labs(title = "Relation between users daily calories burned and Daily Sedentary Hrs",
       subtitle = " 22 Days Data of 24 Fitbit users ")
```
##VIZ.11: Mean Workout Intensity Per Hour in a Day with 31 Days Data of 33 Fitbit Users.
```{r}
ggplot(data = hourlyintensities,aes(x=Time,y=TotalIntensity,fill=DayName)) +
  geom_bar(stat = "summary",fun="mean") +
  theme(axis.text.x = element_text(angle = 90))+
  scale_x_discrete(limits = c("5:00:00 AM","6:00:00 AM","7:00:00 AM",
                              "8:00:00 AM","9:00:00 AM","10:00:00 AM",
                              "11:00:00 AM","12:00:00 PM","1:00:00 PM",
                              "2:00:00 PM","3:00:00 PM","4:00:00 PM",
                              "5:00:00 PM","6:00:00 PM","7:00:00 PM",
                              "8:00:00 PM","9:00:00 PM","10:00:00 PM" )) +
  labs(title = "Mean Workout Intensity Per Hour in a Day",y="MeanTotalIntensity",
       subtitle="31 Days Data of 33Fitbit Users")

```

